---
title: "Dynamicmodel2logistic"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(deSolve)
```

## Difference Equations

Population models can be discrete (rather than continuous)
So we could implement them as difference equations and iterate
like we did with diffusion

```{r iteratie}
source("../R/discrete_logistic_pop.R")
# notice how a for loop is used to iterate

# how many rabbits after 50 years given a growth of 0.1
# starting with 1 rabbit - but a carrying capcity of 500



discrete_logistic_pop(P0=1, r=0.05, K=200, T=50)

# save results
discrete_result = discrete_logistic_pop(P0=1, r=0.05, K=200, T=50)

# lets also keep the parameters for use later
P0=1
r=0.05
K=200
T=50

```

## Differential Equation comparison

Remember we have 3 ways now to calculate population

analytical solution - based on integration (exppop.R) BEST

using an ode solver for numerical approximation (exppop_play.R)

numerical integration using in discrete steps (discrete_logistic_pop.R)
```{r closed}

source("../R/exppop.R")

exppop(P0=P0, r=r, K=K, T=T)
analytic_result = exppop(P0=P0, r=r, K=K, T=T)

analytic_result
discrete_result

# why are they different
# look at trajectories

growth_result = data.frame(time=seq(from=1,to=100))

growth_result$Panalytic = sapply(growth_result$time, exppop, P0=1,r=0.05, K=200)

growth_result$Pdiscrete = sapply(growth_result$time, discrete_logistic_pop, P0=1,r=0.05, K=200)

tmp = growth_result %>% gather(key="Ptype",value="P",-time)
ggplot(tmp, aes(time,P, col=Ptype))+geom_point()

# try running them for longer time periods to see what happens 
# change the value of r, K , P0 - see how it effects the results
```

Finally look at continuous derivative using ODE solve
Needs
*initial condtions
*differential equation
*parameters

```{r}

source("../R/exppop_play.R")

# set up using the same parameters
pcompare = list(r=r,K=K)


# now run our ODE solver
result = ode(P0, growth_result$time, dexppop_play, pcompare)
head(result)
growth_result$Pdifferential = result[,2]

# comapre all 3 approaches
tmp = growth_result %>% gather(key="Ptype",value="P",-time)
ggplot(tmp, aes(time,P, col=Ptype))+geom_point()

# notice Pderiv is closer to Pa
```

All differential and difference equations are approximations
The analytical solution is exact

Notice that differential equations is a bit more accurate!

# Sensitivity Analysis of a differential equation 

We can apply sensitivity analysis to a differential equation

A key issue where is sensitivity of what?

Dyanmics models often give you many many outputs - time series (streamflow every day for a year, popultation for 30 years) - or output over space (concentration every 10 meters)

So if we are asking 'sensitivity of what' we need to summarize results in some way (reduce their dimensionality )

Ideas?


Depends on what is important for your model application

* max

* mean

* min

* total

* variation

So a key step in sensitivity analysis with a dynamics model is summarizing results into a few key measures

Its useful to turn that summarizing activiity into a function


## Workflow

* obtain parameter sets (from sobel of LHS)

* build a function that will extract the information (metrics) you want from your dynamics model (output of the ode)

* create a data structure to store the metrics for each parameter set - lets call it metrics (but could be anything)

* run ODE for each parameter sets to fill in this metrics data structure

* send the metrics data structure back to the sensitivity analysis object (from sobel or LHS)

* plot and analyze results
```{r sen}
source("../R/exppop_play.R")

# lets start with sobel 
library(sensitivity)

# come up with first set of sample parameters
# we will assume that we know the initial population,

Pinitial=10

# want to learn about sensitivity to growth rate (r) and carrying capacity (K)
# set the number of parameters
np=1000
K = rnorm(mean=200, sd=50, n=np)
r = rnorm(mean=0.05, sd=0.005, n=np)
X1 = cbind.data.frame(r=r, K=K)

# repeat to get our second set of samples
K = rnorm(mean=200, sd=50, n=np)
r = rnorm(mean=0.05, sd=0.005, n=np)
X2 = cbind.data.frame(r=r, K=K)

# create our sobel object and get sets ofparameters for running the model

sens_P = sobol2007(model = NULL,X1, X2, nboot = 300)

# our parameter sets are
head(sens_P$X)

# run our differential equation and keep the output
# BUT
# what output do we want  to keep
# how about maximum population if we run the model for 200 years, and how many years to get to the carrying capacity

# for illustration lets look at running just one parameter sets and summarizing results
sens_P$X[1,]
# recall ODE needs ALL of our parameters in a single list 
# initial population and timnes for which we want output 
Pinitial

# gets results for 200 years (evaluating every year)
simtimes = seq(from=1, to=200)
parms = list(r=sens_P$X$r[1], K=sens_P$X$K[1])
result = ode(y=Pinitial, times=simtimes, func=dexppop_play, parms=parms)

head(result)
colnames(result)=c("time","P")
# turn it into a data frame
result = as.data.frame(result)
ggplot(result, aes(time, P))+geom_point()

# extra our metrics of interest  from this
# maximum population it gets to
maxpop = max(result$P)
maxpop

# years required to get to the maximum poputation
# which.max will tell us when max population occur
idx = which.max(result$P)
# turn this index into a year (might be the same if time step in 1 but just in case it isn't)
maxyear = result$time[idx]
maxyear
```

What if we want to run for all parameters

```{r sen2}
# turn computing our metrics into a function

compute_metrics = function(result) {
  maxpop = max(result$P)
idx = which.max(result$P)
maxyear = result$time[idx]
return(list(maxpop=maxpop, maxyear=maxyear))}

# try it on our first parameter set
compute_metrics(result)

# great but we need to apply the ode and this function for all of our parameters



# define a wrapper function to do everything we need - run solver and compute metrics - and send back results for each parameter

p_wrapper = function(r,K, Pinitial, simtimes, func) {
    parms = list(r=r, K=K)
    result = ode(y=Pinitial, times=simtimes, func=dexppop_play, parms=parms) 
    colnames(result)=c("time","P")
  # get metrics
  metrics=compute_metrics(as.data.frame(result))
  return(metrics)
}

# now use mapply as we did before

allresults = mapply(FUN=p_wrapper, r=sens_P$X$r, K=sens_P$X$K,
 MoreArgs=list(Pinitial=Pinitial, simtimes=simtimes, func=dexppop_play))


allres = as.data.frame(matrix(unlist(allresults), ncol=2, byrow=T))
colnames(allres) = c("maxpop","maxyear")

# create boxplots
tmp = allres %>% gather(key="metric", value="value")
ggplot(tmp, aes(metric, value, col=metric))+geom_boxplot()
```

Compute the sobol indicies for each metric

```{r sen3}
# sobol can only handle one output at a time  - so we will need to do them separately

sens_P_maxpop = sensitivity::tell(sens_P,allres$maxpop)

# first-order indices (main effect without co-variance)
sens_P_maxpop$S

# total sensitivity index -note that this partitions the output variance - so values sum to 1
sens_P_maxpop$T

plot(sens_P_maxpop)


# create another one for max year
sens_P_maxyear = sensitivity::tell(sens_P,allres$maxyear)
# first-order indices (main effect without co-variance)
sens_P_maxyear$S

# total sensitivity index -note that this partitions the output variance - so values sum to 1
sens_P_maxyear$T

# notice how K is much more imprtant here
plot(sens_P_maxyear)


```

Assignment



