---
title: "Glue"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(ecohydroexample)
library(tidyverse)
```


```{r setup, echo=FALSE}

load("Data/simQ.rda")
load("Data/simstats.rda")
nparm = nrow(simstats) 
```

Create a single measure of accuracy - we will combine NSE and an estimate of mean percent error
Note that 'count' all measures equally we need to make sure they all go between 0-1, and closer to 1 is a 'better' fit

Larger percent error is 'worse' model fit, so we will transform
(1-abs(Percent.Error)/100), and make any negative values (eg. percent error greater than 100 percent 0)

For NSE we make all values less than 0, 0

We multiply to get a combined accuracy measure

Then use the combined accuacy measure to define behavioural parameter set (**simstats_acc**) - two options

*define a threshold (we will use 30%)
* take top 50 performing parameter sets

(we go with the latter but code could be commented to go with threshold approach)

```{r setup, echo=FALSE}

simstats$accuracy=pmax(0.0,simstats$nse)*pmax(0.0,(1-abs(simstats$perr.total/100)))
# pmax simply takes the maximum of the two values; we do this to remove negative values of performance measures by making them zero
# look at range of accuracy
summary(simstats$accuracy)

# 1) selecting behaviorial or acceptable parameters sets

threshold = 0.3
simstats_acc = subset(simstats, simstats$accuracy > threshold)

# as an alternative  what if you want the top N parameter sets
topN = 50
tmp = simstats[order(simstats$accuracy,decreasing=T),]  
simstats_acc=tmp[1:topN,]

```

Now define "weights" (likelihood) based on parameter performance for the acceptable or behaviorial parameters

We want the sum of the weights to equal 1

* accuracy measure defined above will define weight
* we normalize by the range of accuracy for the behavioural parameters  
* this **relative accuracy ** becomes the weight
* note we now only work with behavioural parameter sets (in ** simstats_acc ** versus ** simstats **)



```{r weighting, echo=FALSE}

# create a weight for each parameter set based on its relative accuracy
max_acc=max(simstats_acc$accuracy)
min_acc=min(simstats_acc$accuracy)
simstats_acc$w_acc=(simstats_acc$accuracy-min_acc)/(max_acc-min_acc)
sum_acc=sum(simstats_acc$accuracy)
simstats_acc$wt_acc=simstats_acc$accuracy/sum_acc

Nacc = nrow(simstats_acc)

```

One way to use weights is to define a maximum likelihood estimate by averaging (weighted by accuracy) streamflow from all behavioural simulations 

We use the **row** in simstats_acc to find the appropriate modelled streamflow in ** simQ **

```{r mle, echo=FALSE}

# generate a streamflow as weighted average of all  acceptable parameter sets

accept_cols = simstats_acc$row
accept_wts = simstats_acc$wt_acc
aver_flow = apply(simQ[,accept_cols],1, weighted.mean, accept_wts)

# add some date information or simply add to simQ

simQ$mleQ = aver_flow

ggplot(simQ, aes(x=date, y=mleQ))+geom_line(col="red")+labs(y="Streamflow mm/day")


```

We could plot all behavioral estimates to range given behavior parameters

Or better, use weights to define upper and lower streamflow quantiles (using **wtd.quantile** from Hmisc package)
We will need the number from the column name to link with weights

```{r plotting, echo=FALSE}

#all parameters
a = simQ[,accept_cols]
b = simQ[,c("date","month","year","day","wy")]
tmp = as.data.frame(cbind(a,b))
Q_acc = tmp%>% gather(key="sim",value="Q", -date,-month,-year,-day,-wy)
ggplot(Q_acc, aes(x=date, y=Q))+geom_line()+theme(legend.position = "none")

# get quantiles for each day
accQ_quantile=apply(simQ[,accept_cols],1,"wtd.quantile",weights=accept_wts,prob=c(0.05,0.5, 0.975),normwt=T)
accQ_quantile=as.data.frame(t(accQ_quantile))
colnames(accQ_quantile)=c("5%","mid", "95%")

accQ_quantile = as.data.frame(cbind(accQ_quantile, simQ[, c("date","month","year","day","wy")]))

accQ_quantilel = accQ_quantile %>% gather(key="quant",value="Q",-date,-month,-year,-day,-wy)

ggplot(accQ_quantilel, aes(x=date, y=Q, col=quant))+geom_line()+labs(y="streamflow (mm/day)")

# just one year
ggplot(subset(accQ_quantilel, wy==1968), aes(x=date, y=Q, col=quant))+geom_line()+labs(y="streamflow (mm/day)")+scale_colour_discrete(name="prediction quantile")

