---
title: "makingfunction3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## User control of functions

Parameters in the function definition can be used to give users options that determine what calculations or actions are done in the function
(model)

This allows functions to be more flexible

Think of a model of ecosystem production that change calculations if vegetation is grass versus a tree


Built in R functions do this alot - they allow users to specify methods (**glm** for example, alows family to set the type of model - binomical or gaussian) 

Here's an example

A function that estimates solar pv power given inputs of radiation

Model inputs: solar radiation (daily direct and diffuse)
Model outputs: power generated each year and average power over time
Parameters: panel efficiency, system performance, units, type of array (uses diffuse or not), plot Y/N

Some of these options such as whether to plot determine outputs from the function
AND the type of array to determine whether it uses diffuse radiation, 
these paramters change how the function/model works

```{r solar}
source("../R/solarpv.R")
solarpv

# read in R formatted data
load("../Data/sierraczosolar.rda")

# already in the formal required for the model
head(sierraczosolar)

# run the model
solarpv(area=0.1, solar=sierraczosolar, clr="green", eunit="W")

# run and save results - but don't plot
site1 = solarpv(area=0.1, solar=sierraczosolar, clr="green", eunit="W", g=FALSE)
site1$mean

# consider a different pv array that only uses 
# diffuse and has non standard efficiency (0.6)
site2 = solarpv(area=0.1, solar=sierraczosolar, clr="green", eunit="W", g=FALSE, eff=0.6, etype="direct")
site2$mean

# try some additional variations
# add an output that is the minimum annual solar output
# add error checking that would remove years that have less than 12 months of data

```

## Building models with blocks - functions that call other functions


Functions can always embed other functions - and in models the output of one model may be the input to another

* loosely coupled models - output from one becomes input the other - but not necessarily automatic
* fully coupled - models exchange input sna outputs (feedback systems)


As an example, compute net present value of power facility
We will use output from our solarpv function as input into a function that
estimate profit given a price - and also 'calls' another function to compute 
net present value

```{r valuep}

source("../R/compute_NPV.R")
source("../R/compute_profit_frompower.R")

compute_profit_frompower

compute_NPV

# use site1 from above as input

profit_solar = compute_profit_frompower(energy=site1$annual$elect, price=30, discount=0.04)
profit_solar

ggplot(profit_solar$yearprofit, aes(year, netpre))+geom_col()+labs(y="Profit in current $", x="time")

# notice that the first
```

Notice that we can 're-use' our function - combining it with
other sources of power data
(such as our reservoir power model)

```{r reuse}

# rerun 'sampling' in makingfunction1.Rmd. above to get results from our reserovir model
head(reservoir_model_res)

# create profit from hydro
profit_hydro = compute_profit_frompower(energy=reservoir_model_res$power, price=30, discount=0.04)

# compare with
profit_solar = compute_profit_frompower(energy=site1$annual$elect, price=30, discount=0.04)

profit_solar$total/1000
profit_hydro$total/1000

```
