---
title: "Marry"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A bit more on data structures

# Multi-dimension arrays

Multi-dimension arrays can be useful when you want to keep track of something that varies in more than 2 ways (rows and columns)


Examples?

Multi-dimensionarray can be useful- but tricky - 
in R sometimes useful for
doing calculations and then we can transform them to data frames for plotting

Imagine
soil moisture at 2 depths for 5 different farms
and 4 different orchard types within each farm

* fruits: 
  avocado, apple,  orange  almond - so 4 fruits
 
* soil moisture 
  store soil moisture at 2 depths 
 ( 25cm, 0.5m)
 
* 5 farms (A through E)


We use the *array* command to define multi-dimensional arrays
and *dim* to check how many dimensions an array has

Also lets us look at nested *for* loops

```{r mdim}

fruit = c("avocado", "apple", "orange",  "almond")
depth = c("25","50")
farm = c("A","B","C","D","E")
soilm = array(dim=c(5,4,2))
dim(soilm)
soilm
# or 

soilm = array(dim=c(length(farm), length(fruit), length(depth)),
              dimnames = list(farm,fruit,depth))
# put some data in the array
for (i in 1:5) {
  for (j in 1:4) {
    value = runif(min=0.2,max=0.5,n=2)
    soilm[i,j,]=value
  }
}
# to see whole array
soilm

# access an element - 3 dimensions so 3 numbers
soilm[1,,]

soilm[,1,]

soilm[,,1]

# how would you get all measurements for the 50cm depth
# how would you get all values for oranges
# how would you get all values for avocadoes in farm C

# single values
soilm[1,2,2]
```

Once we have a multi-dimensional array we can apply calculations over the different dimenstions

We just need to keep track of which dimension is which
Farm- Fruit - Soil

```{r}



# overall whole area and all depths
# easiest!
averagesoilm = mean(soilm)
averagesoilm

#lets say we want to keep all of the locations (farms and fruit) but average 
# over the two depths - average over the 3rd dimension

# note that we leave out the dimension we can to 'get rid of'
averagegrid = apply(soilm, c(1,2),mean)
averagegrid

# now lets say we want to also find the average for each farm
# keep the first dimension (5-farms) but average over others
averagefarm = apply(soilm, c(1), mean)
averagefarm

# average by fruit
averagefruit = apply(soilm, c(2), mean)
averagefruit

# average by  soil moisture depth
averagedepth = apply(soilm, c(3), mean)
averagedepth

# you can use other functions as well
sumfruit = apply(soilm, c(2), sum)
sumfruit

# including your own function
isdrought = function(sm, thresh) {
  n = length(sm)
  drought=TRUE
  for (i in 1:n) {
    if (sm[i] > thresh) drought=FALSE
  }
  return(drought)
}
soilm
#local drought
apply(soilm, c(1,2), isdrought, thresh=0.4)

#farm drought
apply(soilm, c(1), isdrought, thresh=0.5)

#total area drought
isdrought(soilm, thresh=0.3)
isdrought(soilm, thresh=0.7)

```

finally you might not want to keep things in a multi-dimensional array

We can use *as.data.frame.table*

```{r}

soilm_asdataframe = as.data.frame.table(soilm, responseName="Theta")
colnames(soilm_asdataframe)=c("farm","fruit","depth","theta")
ggplot(soilm_asdataframe, aes(farm,theta, col=fruit))+facet_wrap(~depth)+geom_point(size=3)


```

