---
title: "Optimize"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Simple Example

Pollution output as a function of temperature

```{r}
pollutant_export = function(x) {
        result =( (x-2)**2)
				return(result) 
        }

temps = seq(from=-3,to=10)
plot(temps, pollutant_export(temps), col="green", cex=5, pch="*")




```

What temperature minimizes pollutant export?

Easy to see from graph

But can use

*optimize*

requires function to be minimized, lower and upper bounds of parameter, and tolerance (optional)

tolerance is how close do you need to get to minimum

Returns parameter value that gives the minimum
and how close you got

```{r}

result=optimize(pollutant_export, lower=-3, upper=10)
result
```

A more complex example

Lets return to a model of crop yields - we have a new model - in compute_yield that includes the impact of irrigation on yields, given some crop parameters and precipitation and temperature

**compute_yield**

We also have a model of profit that accounts for price of the crop and costs of irrigation

**compute_profit**

What is the optimal irrigation amount given price of crop and cost of irrigation?

We can use **optimize** here - leave one input "open" - e.g use for upper and lower bounds in the optimation
We will keep irrigation "open"
```{r irrigationopt}

source("../R/compute_profit.R")
source("../R/compute_yield.R")
source("../R/compute_NPV.R")

crop_pars = c(ts=0.4, tp=0.5, base.yield=500, Topt=25, max.water=800)

# lets assume water cost $150 per mm/yr (depth of water per crop area)
optimize(compute_profit, lower=0, upper=400, T=c(20,30,25), P=c(250,144,100), discount=0.01, price=50, cost=150, crop.pars=crop_pars,  maximum=TRUE)


# try with cheaper water
optimize(compute_profit, lower=0, upper=400, T=c(20,30,25), P=c(250,144,100), discount=0.01, price=50, cost=25, crop.pars=crop_pars,  maximum=TRUE)


```

You can run your optimization for a range of water prices

```{r optrange}

water_price=seq(from=1, to=50, by=0.5)

# use map to run optimization for different water prices
# lets also broaden our irrigation range
result = water_price %>% map_df(~optimize(compute_profit, lower=0, upper=1000, T=c(20,30,25), P=c(250,144,100), discount=0.01, price=50, cost=.x, crop.pars=crop_pars,  maximum=TRUE))

# add water price to result
result$water_price = water_price

# graph the results
ggplot(result, aes(water_price, maximum))+geom_point(col="green", size=2)
 
# how could you do this for a range of prices for crops

```