---
title: "predprey
"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(deSolve)
library(ggpubr)
library(tidyverse)
library(sensitivity)
library(purrr)
```

Predictor Prey - 

* Ordinary Differential Equation with two dependent variables
* Still use **ODE** solve in R
* use lists to bring in initial conditions for all dependent variables; and outputs derivatives for all dependent variable


Preditor and Prey with Carrying Capacity


```{r examples}

source("../R/lotvmodK.R")

# initial conditions
currpop=c(deer=10, pred=1)

# worried that climate change is going to reduce the carrying capacity (how many deer can be supported, because of drought reducing available foliage for them to eat) -and that this might lead to mountain lion extinction -  some estimates suggested that carrying capacity could be reduced by 50%. 
# this is a challenging problem because there is a lot of uncertainty in how deer/mountain preditor and prey relationships
# Do an analysis that computes mean and minimum deer and mountain lion populations for two scenarios 1) carrying capacity is between 900-1100 deer and 2) carrying capacity is between 400-600 deer. For both scenarios compute mean and minimum preditor and prey estimates with uncertainty bounds! Plot them together and based on this make an assessment of what you think the risk of mountain lion extintion is
# Here are parameters

#Inital deer population (mean 300, standard devidation 20)
#Initial mountain lion population (mean 20, standard deviation 5)

#deer population growth rate (rprey) mean (0.6 standard deviation 0.1)
#mortality rate of mountain lion (pmort) mean (0.5 standard deviation 0.05)
#Efficiency mean (0.8, standard deviation 0.2)
#Alpha mean 0.2 standard deviation 0.05 
eff=0.8
alpha = 0.4
K=100
pmort=0.3
rprey=0.04

prey_i = pmort/(eff*alpha)
pred_i = rprey/alpha * (1-prey_i/K)
# compute results for the mean

currpop=c(prey=prey_i, pred=pred_i)

currpop = c(prey=800, pred=1)
pars = c(rprey=rprey, alpha=alpha, eff=eff,pmort=pmort, K=K)

# times when you want to evaluate
years = seq(from=1,to=5, by=0.1)

# run our differential equation solver
res = ode(func=lotvmodK, y=currpop, times=years, parms=pars, method="euler")

# estract the results
ressimple = as.data.frame(res) %>% gather(key="animal", value="pop",-time)

# graph both populations over time
p1=ggplot(ressimple, aes(time, pop, col=animal))+geom_line()
p1

# also look at relationships between preditor and prey population and use color for time 
# I will remove the legend here to make it easier to see 
p2 = ggplot(as.data.frame(res), aes(pred, prey, col=as.factor(round(time/10))))+geom_point()+theme(legend.position = "none")
p2

ggarrange(p1,p2)

# try with different parameter sets, can you create one where populations are stable - less cycling?

```


# Sensitivity analysis

Consider pred-prey 
BUT what will be the output - if we want to 'quantify sensitivity'
useful to look at a single value or set of value

for example
Max Prey Pop
Min Prey Pop


Remember steps

* Generate parameters (LHS, Sobel)
* Metrics function
* Wrapper Function
* Run wrapper function to get metrics for all paramater sets
* Graph and compute sensitivity statistics

```{r odesen}
source("../R/lotvmodK.R")
library(pse)

# create parameter information as above
factors = c("rprey","K","alpha","eff","pmort")
nsets=500
q = c("qunif","qunif","qunif","qnorm", "qunif","qunif","qunif")
q.arg = list(list(min=0.01, max=0.3), list(min=10, max=200), list(min=0.1, max=0.4), list(mean=0.3, sd=0.01), list(min=0.05, max=0.45))
sens_pp = pse::LHS(NULL,factors,nsets,q,q.arg)
sens_pars = get.data(sens_pp)


# lets create a metric and wrapper function as we did for our population models

# first our metrics 
# lets say we  want the maximum and minimum  of both predictor and prey

compute_metrics = function(result) {
  maxprey = max(result$prey)
  maxpred = max(result$pred)
  minprey = min(result$prey)
  minpred = min(result$pred)
return(list(maxprey=maxprey, minprey=minprey, maxpred=maxpred, minpred=minpred))}

# build a wrapper function


p_wrapper = function(rprey,alpha, eff, pmort, K, currpop, days, func) {
    parms = list(rprey=rprey, alpha=alpha, eff=eff, pmort=pmort, K=K)
    result = ode(y=currpop, times=days, func=func, parms=parms) 
    colnames(result)=c("time","prey","pred")
  # get metrics
  metrics=compute_metrics(as.data.frame(result))
  return(metrics)
}


# run our model for all parameters and extract the results
currpop=c(prey=1, pred=1)
days = seq(from=1,to=500)
allresults = sens_pars %>% pmap(p_wrapper, currpop=currpop, days=days, func=lotvmodK)
 
# take results back to unlisted form
allres = allresults %>% map_dfr(`[`,c("maxprey","minprey","maxpred","minpred"))


# range of response across parameter uncertainty
allresl = allres %>% gather(key="metric",value="pop")
ggplot(allresl, aes(metric, pop))+geom_boxplot()

# dealing with different scales
ggplot(allresl, aes(metric, pop, col=metric))+geom_boxplot()+facet_wrap(~metric, scales="free")

# send results back to our sensitivity analysis data structure
sens_pp = pse::tell(sens_pp, t(allres), res.names=c("maxprey", "minprey", "maxpred","minpred"))

# now look at sensitivity
# cummulative distributions
pse::plotecdf(sens_pp)

# parameter sensitivity
plotscatter(sens_pp)

# metrics of sensitivity
plotprcc(sens_pp)
sens_pp$prcc
```



How do we think about stablity?

Populations don't change when derivatives are zero!

What conditions lead to BOTH derivatives being zero

For lotvmod
* dprey = rprey*prey -  alpha*prey*pred
* dpred = eff*alpha*prey*pred - pmort*pred

Make dprey and dpred equal to 0 and rearrange
* pred_i = rprey/alpha
* prey_i = pmort/(eff * alpha)

For lotvmodK

* dprey = rprey*(1-prey/K)*prey -  alpha*prey*pred
* dpred = eff*alpha*prey*pred - pmort*pred

Make dprey and dpred equal to 0 and rearrange

* prey_i = pmort/(eff*alpha)
* pred_i = rprey/alpha * (1-prey_i/K)
Try setting you initial conditions close to these values and see what happens


```{r stabilitynew}

# set parameter list
pars = data.frame(rprey=0.1, alpha=0.6, eff=0.8,pmort=0.4, K=20)

# now lets try initial conditions that will be stable
preyi = with(pars, pmort/(eff*alpha))
predi = with(pars, rprey/alpha*(1-preyi/K))
# times when you want to evaluate
days = seq(from=1,to=500)

# lets first see what happens when we start with 1 of each
currpop=c(prey=1, pred=1)
# run our differential equation solver
res = ode(func=lotvmodK, y=currpop, times=days, parms=pars)
# extract the results
res_smallstart = as.data.frame(res) %>% gather(key="animal", value="pop",-time)
# graph both populations over time
p1=ggplot(res_smallstart, aes(time, pop, col=animal))+geom_line()
p1

# lets first see what happens when we start our estimates of stable populations
stablepop = c(prey=preyi, pred=predi)
res = ode(func=lotvmodK, y=stablepop, times=days, parms=pars)
# estract the results
res_stablestart = as.data.frame(res) %>% gather(key="animal", value="pop",-time)
# graph both populations over time
p2=ggplot(res_stablestart, aes(time, pop, col=animal))+geom_line()
p2

# of course in this case these are not very realistic populations, so it is unlikely that you would get to something stable given parameters - but if they were different you might find a stable and realistic population

# try
```
And just for fun

Lets look at a Lorenz System with its interesting dynamics
```{r lorenze}
# lorenze
source("../R/lorenz.R")
pars = list(a=10,b=28,c=8/3)
res = ode(func=lorenz, c(x=0.1,y=0,z=0), times=seq(0,50,by=0.01), parms=pars)

ggplot(as.data.frame(res), aes(x,y, col=time))+geom_point()
ggplot(as.data.frame(res), aes(x,z, col=time))+geom_point()
ggplot(as.data.frame(res), aes(y,z, col=time))+geom_point()

ressimple = as.data.frame(res) %>% gather(key="var", value="value",-time)
ggplot(ressimple, aes(time, value, col=var))+geom_line()

# try with different initial conditions
pars = list(a=15,b=28,c=8/4)
res = ode(func=lorenz, c(x=0.3,y=5,z=10), times=seq(0,50,by=0.01), parms=pars)

ggplot(as.data.frame(res), aes(x,y, col=time))+geom_point()+scale_colour_gradientn(colours = terrain.colors(10))
ggplot(as.data.frame(res), aes(x,z, col=time))+geom_point()+scale_colour_gradientn(colours = terrain.colors(10))
ggplot(as.data.frame(res), aes(y,z, col=time))+geom_point()+scale_colour_gradientn(colours = terrain.colors(10))

ressimple = as.data.frame(res) %>% gather(key="var", value="value",-time)
ggplot(ressimple, aes(time, value, col=var))+geom_line()


```

